### 2014-09-22 stat-table.tex

% latex table generated in R 3.1.1 by xtable 1.7-3 package
% Fri Sep 19 11:41:03 2014
\begin{table}[ht]
\centering
\begin{tabular}{lrrrrrrr}
  \hline
taxonomic\_class & n & n\_orders & n\_species & median\_N & max\_N & n\_interpolated & n\_zero\_sub \\ 
  \hline
Aves & 227 &  15 & 113 &  26 & 117 &  68 &  32 \\ 
  Insecta & 182 &   7 &  91 &  25 &  87 &  26 &  55 \\ 
  Mammalia & 125 &   8 &  51 &  28 & 106 &  18 &  21 \\ 
  Osteichthyes & 109 &   6 &  36 &  26 &  67 &  14 &   4 \\ 
  Crustacea &  13 &   2 &   2 &  28 &  33 &   1 &   0 \\ 
  Chondrichtyhes &   1 &   1 &   1 &  20 &  20 &   1 &   0 \\ 
  Gastropoda &   1 &   1 &   1 &  21 &  21 &   0 &   0 \\ 
   \hline
\end{tabular}
\end{table}