### 2014-09-25 stat-table.tex

% latex table generated in R 3.1.1 by xtable 1.7-3 package
% Thu Sep 25 13:22:15 2014
\begin{table}[ht]
\centering
\caption{Summary statistics for the filtered Global Population Dynamics Database time series arranged by taxonomic class. Columns are: number of populations, number of taxonomic orders, numbers of species, median time series length, total number of interpolated time steps, and total number of substituted zeros.} 
\begin{tabular}{lrrrrrr}
  \toprule
Taxonomic class & Populations & Orders & Species & Median length & Interpolated pts & Zeros pts \\ 
  \midrule
Aves & 227 &  15 & 113 &  26 &  68 &  32 \\ 
  Insecta & 182 &   7 &  91 &  25 &  26 &  55 \\ 
  Mammalia & 125 &   8 &  51 &  28 &  18 &  21 \\ 
  Osteichthyes & 109 &   6 &  36 &  26 &  14 &   4 \\ 
  Crustacea &  13 &   2 &   2 &  28 &   1 &   0 \\ 
  Chondrichtyhes &   1 &   1 &   1 &  20 &   1 &   0 \\ 
  Gastropoda &   1 &   1 &   1 &  21 &   0 &   0 \\ 
   \bottomrule
\end{tabular}
\end{table}