### 2014-10-02 anderson-etal-blackswan-timeseries.tex
Switch to markdown ms; add to ms

\documentclass[12pt]{article}
\usepackage{geometry}
\geometry{verbose,letterpaper,tmargin=2.54cm,bmargin=2.54cm,lmargin=2.54cm,rmargin=2.54cm}
\geometry{letterpaper}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{lineno}
\usepackage{ragged2e}
\usepackage{color}

\usepackage{longtable}
\usepackage{multirow}
\usepackage{array}
\usepackage{tabu} % for spacing between rows in longtable
\setlength{\tabulinesep}{9pt}

 % Linux Libertine:
  \usepackage{textcomp}
     \usepackage[sb]{libertine}
     \usepackage[varqu,varl]{inconsolata}% sans serif typewriter
     \usepackage[libertine,bigdelims,vvarbb]{newtxmath} % bb from STIX
     \usepackage[cal=boondoxo]{mathalfa} % mathcal
     \useosf % osf for text, not math
     \usepackage[supstfm=libertinesups,%
       supscaled=1.2,%
       raised=-.13em]{superiors}

\mathchardef\mhyphen="2D

\textheight 22.0cm

\usepackage[round]{natbib}
\bibliographystyle{apalike}
\bibpunct{(}{)}{;}{a}{}{;}

\title{Black swans in ecological time series\\Black swans in population dynamics\\Heavy-tailed ecological time series\\Black swans in ecological dynamics}
\author{
Sean C. Anderson$^{1\ast}$ \and
Trevor A. Branch$^2$ \and
Andrew B. Cooper$^3$ \and
Nicholas K. Dulvy$^1$
}
\date{}

% remove numbers in front of sections:
\makeatletter
\renewcommand\@seccntformat[1]{}
\makeatother

\setlength\parskip{0.16in}
\setlength\parindent{0in}
\usepackage{titlesec}
\titlespacing\section{0pt}{12pt plus 4pt minus 2pt}{-6pt plus 2pt minus 2pt}
\titlespacing\subsection{0pt}{12pt plus 4pt minus 2pt}{-8pt plus 2pt minus 2pt}

\begin{document}
\input{values} % R output

%\begin{spacing}{1.4}
%\doublespacing
\onehalfspacing

\maketitle
\thispagestyle{empty}

\input{ms}

\bibliography{/Users/seananderson/Dropbox/tex/jshort,/Users/seananderson/Dropbox/tex/ref3}
%\bibliography{/Users/seananderson/Dropbox/tex/jshort,refs}

\clearpage
\section{Tables}

\input{sparks.tex}

\clearpage
\section{Figures}

\input{figs.tex}

\begin{centering}
\LARGE
Supporting material\\[1.5em]
\end{centering}

\input{supp}

\input{supp-figs}

\end{document}