### 2014-10-09 anderson-etal-blackswan-timeseries.tex
A whole lot of paper writing/editing

\documentclass[12pt]{article}
\usepackage{geometry}
\geometry{verbose,letterpaper,tmargin=2.54cm,bmargin=2.54cm,lmargin=2.54cm,rmargin=2.54cm}
\geometry{letterpaper}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{lineno}
\usepackage{ragged2e}
\usepackage{color}
\usepackage{verbatim}
\usepackage[sectionbib]{chapterbib}

\usepackage{amsmath}
% \makeatletter
% \g@addto@macro\normalsize{%
%   \setlength\abovedisplayskip{-20pt}
%   \setlength\belowdisplayskip{0pt}
%   \setlength\abovedisplayshortskip{-20pt}
%   \setlength\belowdisplayshortskip{0pt}
% }
% \makeatother

\usepackage{longtable}
\usepackage{multirow}
\usepackage{array}
\usepackage{tabu} % for spacing between rows in longtable
\setlength{\tabulinesep}{9pt}

 % Linux Libertine:
  \usepackage{textcomp}
     \usepackage[sb]{libertine}
     \usepackage[varqu,varl]{inconsolata}% sans serif typewriter
     \usepackage[libertine,bigdelims,vvarbb]{newtxmath} % bb from STIX
     \usepackage[cal=boondoxo]{mathalfa} % mathcal
     \useosf % osf for text, not math
     \usepackage[supstfm=libertinesups,%
       supscaled=1.2,%
       raised=-.13em]{superiors}

\mathchardef\mhyphen="2D

\textheight 22.0cm

\usepackage[round,sectionbib]{natbib}
\bibpunct{(}{)}{;}{a}{}{;}

\title{Black swans in ecological time series\\Black swans in population dynamics\\Heavy-tailed ecological time series\\Black swans in ecological dynamics}
\author{
Sean C. Anderson$^{1\ast}$ \and
Trevor A. Branch$^2$ \and
Andrew B. Cooper$^3$ \and
Nicholas K. Dulvy$^1$
}
\date{}

% remove numbers in front of sections:
\makeatletter
\renewcommand\@seccntformat[1]{}
\makeatother

\setlength\parskip{0.16in}
\setlength\parindent{0in}
\usepackage{titlesec}
\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{-6pt plus 2pt minus 2pt}
\titlespacing\subsection{0pt}{0pt plus 4pt minus 2pt}{-8pt plus 2pt minus 2pt}

\begin{document}

\maketitle

\input{values} % R output

\emph{Authors are listed alphabetically (ABCD!) and open to reordering.}

\textsuperscript{1}Earth to ocean research group, department of biological sciences, Simon Fraser University, Burnaby BC, V5A 1S6, Canada

\textsuperscript{2}School of Aquatic and Fishery Sciences, University of Washington, Box 355020, Seattle, WA 98195, USA

\textsuperscript{3}School of Resource and Environmental Management, Simon Fraser University, Burnaby, BC, V5A 1S6, Canada

\textsuperscript{*}Corresponding author: Sean C. Anderson; Earth to Ocean Research Group, Department of Biological Sciences, Simon Fraser University, Burnaby BC, V5A 1S6; Phone: 1-778-782-3989; E-mail: sean\_anderson@sfu.ca

%\begin{spacing}{1.4}
\doublespacing
%\onehalfspacing

\thispagestyle{empty}

\linenumbers

\include{ms}

\clearpage
\section{Figures}

\input{figs}

\include{supp}

\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0}

\include{stat-table}

\include{supp-figs}

\end{document}