### 2014-10-18 ms.tex
Wrap ms and (very) minor edits

%!TEX root = anderson-etal-blackswan-timeseries.tex

\noindent
\emph{Main message, to be deleted (25 words)}: Black swans are present \&
taxonomically widespread but rare in population dynamics. Extreme climate,
predation, parasites, and their interactions are common causes; intrinsic
drivers are unclear.

\section{Abstract}

Black swans are statistically improbable events that nonetheless occur --- often
with profound implications. While extremes in the physical environment,such as
monsoons and heat waves, are widely studied and increasing in magnitude and
frequency, it remains unclear the extent to which ecological populations buffer
or suffer from extremes. Here, we develop a probability model to estimate the
degree of heavy-tailedness (presence of black swans) in ecological process
noise. We apply our model to \NPops~time series from around the world across
\NOrders~taxonomic orders and seven classes. We find strong evidence of black
swans, but they are rare, occurring in \overallMinPerc--\overallMaxPerc\%of
populations; most frequently for birds (\AvesRangePerc\%) followed by mammals
(\MammaliaRangePerc\%), insects (\InsectaRangePerc\%), and
fishes(\OsteichthyesRangePerc\%). When they occur, they tend to be driven by
climate and severe winters, cycles of parasites and predators, and interactions
between these elements. They are more frequently detected for populations with
longer time series and lower levels of process noise. We find little evidence
of intrinsic life-history covariates. TODO close with modelling and policy
implications

\section{Introduction}

Black swans are unexpected extreme events with dramatic consequences. One of
the most striking black swans in ecology was the asteroid marking the mass
extinctions of the K-T boundary (REF). Today, it is the extremeness of climate
--- particularly in concert with shifts in mean temperature --- that is
expected to cause the greatest societal damage (REF).

However, while extremes in the physical environment are present and widely
accepted, it remains unclear the extent to which ecological systems buffer or
suffer from black swans.

Existing evidence for black swans or heavy tails in population dynamics is thus
far limited. Considerable work has focussed on heavy tails in physical
environmental processes in, for example, rain fall (REF), wave height (REF),
earth quake magnitude, climate temperature (REF). time to extinction, plankton
time series, GPDD work found little evidence subsequent work suggested, but...
Ward applied a data-intense method and found evidence for...

There are two key reasons why we may find little evidence of ecological black
swans. First, they might not exist. Indeed, the majority of population
dynamics model fitting and risk forecasting assumes that population dynamics
are not heavy tailed (e.g.~REF, REF, REF). Alternatively, population dynamics
might be heavy tailed, but a paucity of statistical methods and shortness of
available time series has limited our ability to detect them. For example,
a commonly applied probability distribution to detect the shape of time series
tails (the generalized extreme value distribution) requires lumping data into
the most extreme event per unit time (e.g.~heaviest rainfall per year) and
therefore requires a considerable quantity of data (REF). Previous work into
the shape of population dynamic tails, has focused on model comparison to
identify parsimonious explanations from a set of particular hypotheses, but
does not quantify the magnitude of heavy tails on a continuous scale or put the
conclusions in a probabilistic framework (REF). Furthermore, information
criteria are poor at distinguishing subtle differences in population dynamics
for the relatively short time series that are most readily available (REF CHANGE
THIS?).

Or bring back in Ward - data intensive, can't be applied to most individual
time series

plankton --- hideous distribution to fit, even harder in population dynamics
framework -and what about seasonal cycles, density dependence?

GPDD - model competition and finding most parsimonious explanation from set of
hypotheses --- but only \emph{very} heavy from not heavy

GEV --- super data intensive

tipping point and regime shift methods --- say something\ldots{} not
appropriate for most of the data available

Here we develop a probabilist framework for identifying ecological black swans
in population dynamic process noise --- the stochastic jumps from time step to
time step. Our framework quantifies both the magnitude and probability of
heavy-tailed dynamics, allows for a range of population dynamic models, can
incorporate observation uncertainty, and is applicable to commonly available
lengths of abundance time series. We apply our model to hundreds of populations
from around the world across four taxonomic classes to address three questions:
(1) how frequent and strong are black swans in population dynamics across
taxonomy, (2) what are some verified causes of black swans in ecological time
series, and (3) are there characteristics of time series or intrinsic
life-history characteristics associated with heavy-tailed dynamics. We find
a high probability of ecological black swans across a wide range of taxa but in
a limited number of populations. When they occurred, extreme climate,
predation, parasites, and their interactions were common documented causes. We
found little evidence that intrinsic life-history characteristics were related
to the probability of black swans.

Papers to work in: \citep{inchausti2002, halley2002, inchausti2001}
\citep{jentsch2007} \citep{ward2007} \citep{garcia-carreras2011}
\citep{sornette2009} \citep{nunez2012} \citep{thompson2013}
\citep{beaugrand2012} \citep{pine-iii2009} \citep{doak2008} Extreme rain and
monsoons over India increasing \citep{goswami2006} \citep{smale2013}
\citep{easterling2000} \citep{scheffer2003} \citep{katz2005} \citep{taleb2007}
\citep{vasseur2014} \citep{vert-pre2013} \citep{lindenmayer2010}
\citep{valpine2002} \citep{gregory2010} \citet{garcia-carreras2011}
\citeauthor{brook2006} \citetext{\citeyear{brook2006};
  \citealp{herrandoprez2014}} \citep{sibly2005, ziebarth2010}

\section{Methods}

To obtain estimates of the probability and magnitude of black swans, we fitted
population dynamic models to hundreds of abundance time series from around the
world. The key addition of our models, compared to typical population models,
was estimating the shape of the process noise tails. We then compared how these
estimates related to taxonomy, time series properties, and life-history
characteristics, and for a subset of populations, we documented the established
causes of ecological black swans.

\subsection{Time-series data}

We obtained abundance time series from the Global Population Dynamics Database
(GPDD) \citep{gpdd2010}. The GPDD contains nearly 5000 time series of abundance
across $\sim$1000 species and $\sim$100 taxonomic orders. To derive
a high-quality subset of populations that were suitable for our analyses, we
filtered the data (Supporting Material). Our filtering removed populations from
less reliable data sources, removed those without sufficient data for our
models, and interpolated some missing values. Our final dataset contained
\NPops~populations across \NOrders~taxonomic orders and seven taxonomic classes
with a median of XX time steps (range of XX--XX) (Table S1,
Fig.~\ref{fig:all-ts}).

\subsection{Population models}

Our main analysis focuses on the commonly applied phenomenological Gompertz
population dynamics model. In previous analyses of the GPDD, the Gompertz has
proved to be the most frequently parsimonious model \citep{brook2006}. We
extended the traditional Gompertz model by allowing the process noise to be
drawn from a Student-t distribution with degrees of freedom parameter $\nu$
(Fig.~\ref{fig:didactic}). If $\nu$ is small ($\lesssim 10$) the t distribution
has much heavier tails than a normal distribution. For example, at $\nu = 2$,
the probability of drawing a value more than five standard deviations below the
mean is $1.8$\%, whereas the probability of drawing such a value from a normal
distribution is nearly zero ($2.9\cdot10^{-5}$\%). As $\nu$ approaches infinity
the distribution approaches the normal distribution. By estimating the value of
$\nu$ we can quantity how heavy-tailed the process noise deviations are
(Fig.~\ref{fig:didactic}).

The heavy-tailed Gompertz model is:
\begin{align*}
x_t &= \ln N_t\\
x_t &= \lambda + b x_{t-1} + \epsilon_t\\
\epsilon_t &\sim \mathrm{Student\mhyphen t}(\nu, 0, \sigma),
\end{align*}
where $x_t$ is the $\ln$ abundance ($N$) at time $t$. The model is density
independent if $b = 1$, maximally density dependent if $b = 0$, and inversely
density dependent if $b < 0$. The parameter $\lambda$ represents the expected
$\ln$ number of new individuals at time $t + 1$ if $N_t = 1$ (per capita
population growth rate). The process noise is modelled as a Student-t
distribution with degrees of freedom parameter $\nu$, a mean of $0$, and
a scale parameter of $\sigma$.

We chose weakly informative priors to incorporate our understanding of
plausible population dynamics \citep[Supporting Material]{gelman2014}. For
$\nu$, we chose an exponential prior with rate parameter of $0.01$ truncated at
values above two (Fig.~\ref{fig:priors}), a slightly less informative prior
than that suggested by \citet{fernandez1998}. This prior gives only a 7.7\%
probability that $\nu < 10$ but constrains the sampling sufficiently to avoid
wandering off towards infinity --- above approximately $\nu = 20$ the
t distribution is so similar to the normal distribution that time series of the
length considered here are unlikely to be informative about the precise value
of $\nu$ (Fig.~\ref{fig:didactic}).

We estimated all models in a Bayesian framework using Stan
\citep{stan-manual2014}. Stan samples from the posterior distribution with
a version of Hamiltonian Markov chain Monte Carlo called the No-U-Turn Sampler
\citep{hoffman2014}. We assured that chains had sufficiently converged and the
sampler had obtained sufficient independent samples from the posterior
($\widehat{R} < 1.05$, $n_\mathrm{eff} > 200$; Supporting Material). We used
simulated data to test how easily we could detect $\nu$ given different sample
sizes and to ensure we could recover unbiased parameter estimates from the
Gompertz model (Supporting Material, Fig.~\ref{fig:sim-nu},
\ref{fig:sim-gompertz}, \ref{fig:sim-gompertz-boxplots}).

We fit four alternative population models to test if they systematically
changed our conclusions. (1) Autocorrelation has been suggested as a reason for
increased observed variability of abundance time series through time, which
could create apparent heavy tails \citep{inchausti2002}; therefore, we fit
a model that modelled the serial correlation in the residuals. (2) Previous
work has focussed modelled growth rates without accounting for density
dependence \citep{segura2013}; therefore, we fit a simpler model in which we
assumed density independence. (3) Observation error could bias parameter
estimates \citep{knape2012} or mask our ability to detect heavy tails
\citep{ward2007}. However, simultaneously estimating observation error and
process noise in a state space framework is a challenging computational problem
\citep[e.g.][]{dennis2006, knape2008} even with longer time series and without
estimating the shape of the process error tails; therefore, we fit a model
where we assumed a level of observation error ($0.2$ standard deviations on
a log scale). (4) The Gompertz model assumes that population growth rate
declines linearly with log abundance. We also fit an alternative model, the
Ricker-logistic model, which assumes that population growth rate declines
linearly with abundance itself (Supporting Material).

\subsection{Covariates of population dynamic black swans}

We investigated possible covariates of heavy-tailed population dynamics
visually and through multilevel modelling. We plotted characteristics of the
time series ($\sigma$, $\lambda$, $b$, and time-series length) along with two
life-history characteristics (body length and maximum lifespan obtained from
\citet{brook2006a}) against our estimates of $\nu$. We formally investigated
these relationships by fitting beta regression multilevel models to the
probability that $\nu$ was less than $10$ (i.e.~approximately the probability
of heavy tails). For covariates that were derived from Gompertz model parameter
estimates, we incorporated standard deviations around the means. To account for
broad patterns of phylogenetic relatedness, we allowed for varying intercepts
at the taxonomic class, order, and species level (Supporting Material).

Finally, we investigated a sample of populations that our method categorized as
having a high probability of heavy tails (Pr$(\nu < 10) > 0.5$). Where
possible, we found the documented causes of ecological black swans in the
primary data source cited in the GPDD or in other literature describing the
population. These populations were chosen haphazardly with the purpose of
generating hypotheses about the causes of population dynamic black swans.

\section{Results}

We found strong but rare evidence for black-swan dynamics. Defining black-swan
dynamics as a greater than $0.5$ probability that $\nu < 10$, we observed black
swans most frequently for birds (X\%) followed by mammals (X\%), insects (X\%),
and fishes (X\%) (Fig.~\ref{fig:nu-coefs}, \ref{fig:heavy-ts}). Black swans
were taxonomically widespread, occurring for at least one population in X\% of
the taxonomic orders recorded. Accounting for time series length and partially
pooling inference across taxonomic class and order with a multilevel model,
there was stronger evidence for black swans in insect populations than is
apparent in Fig.~\ref{fig:nu-coefs} --- four of 10 orders with the highest
median probability of heavy tails were insect orders --- however, there was
considerable uncertainty in these estimates (Fig.~\ref{fig:order-estimates}).

The majority of our heavy-tailed estimates were robust to alternative
population models and observation error (Fig.~\ref{fig:alt}). Including an
autocorrelation structure in the residuals, modelling population growth rates
as a random walk with drift, or modelling the population dynamics as
Ricker-logistic did not systematically alter our conclusions. Allowing for
a fixed quantity of observation error decreased our estimates of black-swan
dynamics, increasing the median estimate of $\nu$ from $<10$ to $>10$ in X of
XX populations although the majority of $\nu$ estimates remained qualitatively
similar (Fig.~\ref{fig:alt})

Across populations, the probability of observing black swan dynamics was
positively related to time-series length and negatively related to magnitude of
process noise ($\sigma$) but not clearly related to population growth rate
($\lambda$), density dependence ($b$), or maximum lifespan
(Fig.~\ref{fig:correlates}, \ref{fig:correlate-coefs}). Longer time-series
length was the strongest covariate of observing black swan dynamics (as was
similarly observed in simulation testing, Fig.~\ref{fig:sim-nu}). For example,
we were about 15\% TODO more likely to observe heavy tails in a population with
40 time steps than in one with 30 time steps (Fig.~\ref{fig:perc-inc-p}).
However, the absolute change in probability with increased time series length
was small ($0.X$ vs. $0.X$ in the previous example, Fig.~\ref{fig:correlates}).

We assembled the documented causes of black swan population dynamics in seven
populations with high probabilities of heavy tails (Table 1). The majority of
documented events were downward black swans and the majority involved
a confluence of events. For example, a shortage of nest sites for the shag in
the UK was thought to reduce population productivity and make the population
vulnerable when a red-tide hit in 1968. The population experienced a black swan
downswing, but the new availability of nest sites resulted in a rapid
population upswing \citep{potts1980}. As another example, an interaction
between population cycles caused by predation combined with cycles caused by
the environment are thought to have caused a downward black swan for a water
vole population \citep{saucy1994}. Other black swans were the result of
a sequence of extreme climate events on their own. For example, severe winters
in 1929, 1940--1942, and 1962--1963 were associated with black swan downswings
in grey heron abundance in the UK \citep{stafford1971}. Our analysis finds that
the last event was a combination of two black swan events in a row and
population recovery took three times longer than predicted
\citep{stafford1971}.

\section{Discussion}

We found strong evidence that the multiplicative process noise jumps from time
step to time step were greater than we would expect under a normal distribution
and a Gompertz or Ricker-logistic model in XX--XX\% of populations assessed. We
detected black swans more frequently for longer time series and for populations
with a smaller magnitude of process error, as we might expect. However, we
failed to find strong evidence that black swans were associated with density
dependence, population growth rate, or lifespan. In cases that we verified,
black swans were often a result of the interaction between elements of extreme
climate, predators and parasite cycles, and strong changes in human pressures.
Our empirical results, sensitivity tests, and simulation tests suggest that
estimating the tail shape of process error is a viable method of detecting
heavy-tailed population dynamics. The presence of black swans highlights the
importance of developing management strategies that detect quickly, respond to,
and are robust to extremes in population dynamics --- particularly as the
frequency and magnitude of climatic extremes increase over the next century
(REF).

Our results clarify, but do not disagree with previous related analyses.
Previous work with an older version of the GPDD assessed heavy-tailed behaviour
but focused on identifying if the lognormal distribution was the most
frequently parsimonious model of population abundance \citep{halley2002}. For
heavy-tailed distributions, \citet{halley2002} fit the extremely heavy-tailed
Cauchy distribution and the four-parameter Levy stable distribution and found
little information criteria support for these distributions in longer time
series. However, black swan events are by definition rare, and the majority of
time series in the GPDD are short for these purposes. Therefore, we would not
expect to observe black swans in a large proportion of populations.
Furthermore, by probabilistically quantifying process noise tail behaviour on
a continuous spectrum, and allowing for non-stationary time series and density
dependence, our analysis allows for a more nuanced description of the evidence
of ecological black swans. In an earlier study, \citet{keitt1998} described
heavy (power law) tails in breeding bird population abundance. However, this
finding was challenged by \citet{allen2001}, who showed that mixing data across
species could create falsely generate heavy tails. This rebuttal supports our
use of the t distribution to represent heavy tails, since a t distribution can
be represented by a mixture of normal distributions \citep[with
inverse-gamma-distributed variances,][]{gelman2014}.

We might expect to observe black swan dynamics in ecological time series
because of unmodelled intrinsic properties of populations or extrinsic forces
acting on populations. Since a t-distribution can be formed by a mixture of
normal distributions, we could observe heavy tails if we miss some underlying
mixture of intrinsic processes \citep{allen2001}. That process might be an
aggregation of populations across space, or population diversity, or some
intrinsic change in population variability through time. Extrinsic forces could
also cause black swan dynamics (REF). These forces could be extreme themselves.
For example, extreme climate, predation from or competition with other species
experiencing black swans, or sharp changes in human pressure such hunting,
fishing, or habitat destruction might cause black swans (REFS). Alternatively,
the interaction of multiple ``normal'' extrinsic forces could give rise to
black swan ecological dynamics. This could occur if the interaction nature is
synergistic \citep[e.g.][]{kirby2009} or even if non-synergistic interactions
experience a rare alignment \citep{denny2009}.

There a number of caveats when considering the generality of our results.
First, the GPDD data represent a taxonomically and geographically biased sample
of populations. Also, the longer time series we focus on are dominated by
commercially and recreationally important species and a disproportionate number
of populations are located in the United Kingdom. Although we expect that we
would find strong but rare evidence for black swans in other large taxonomic or
geographic samples, the common forces driving those black swans (e.g.~severe
winters in Table~1) likely differ. Second, in a large database with disparate
data sources, we cannot ensure the quality of every data point. Recording
mistakes undoubtedly comprise some of the black swans we detected. Conversely,
recording mistakes could mask heavy tails if initial observations are discarded
or altered as assumed recording errors. Third, the temporal scale of observation
and population dynamics vary considerably across populations in the GPDD and
this likely influences the detection of heavy tails. For example, if we make
frequent observations relative to generation time (e.g.~for many large bodied
mammals) we will average across generations and perhaps not observe the tails.
Furthermore, if we record population censuses infrequently relative to
generation time (e.g.~many insects in the GPDD) the recorded data may average
across extreme and less-extreme events and dampen observed black swan dynamics.

Our results suggest a number of important future issues to address regarding
ecological black swans. First, given that black swans can and do occur, can we
forecast the probability of black swans in space and time? Furthermore, what
management policies allow us to detect them quickly after they happen?
Secondly, can we isolate the components of ecological dynamics that experience
black swans events by moving from phenomenological models such as the Gompertz
to mechanistic models, that take into account, for example, recruitment? We
expect that greater insight into the mechanisms and covariates of ecological
black swans may be best obtained through specific geographic and taxonomic
subsets of data where longer time series are available
\citep[e.g.][]{segura2013}. Finally, if a sizeable fraction of populations
experience heavy tails in population process noise, what is the impact of
allowing for black swans in forecasts of ecological risk? Financial analysts
are acutely concerned with the shape of the tail distribution of financial
returns since these impact estimates of risk --- the probability of a specific
magnitude of undesired event occurring (REF). It is possible that a comparable
focus in ecology would alter our perception of ecological forecasts such as
extinction risk.

Policy implications

\section{Acknowledgements}

We thank members of the Earth to Ocean research group for helpful discussions
and comments. We are grateful to the contributors and maintainers of the Global
Population Dynamics Database and to Compute Canada's WestGrid high-performance
computing resources. Three silhouette images were obtained from
\texttt{phylopic.org} under a Creative Commons Attribution 3.0 Unported
license: rabbit (Sarah Werning), hoverfly (Gareth Monger), bird (Jean-Raphaël
Guillaumin {[}photography{]} and T. Michael Keesey {[}vectorization{]}).
Funding was provided by a Simon Fraser University Graduate Fellowship (SCA),
the Natural Sciences and Engineering Research Council of Canada (NKD, ABC), the
Canada Research Chairs Program (NKD), and XX-TODO (TAB).

\bibliographystyle{ecologyletters}
\bibliography{/Users/seananderson/Dropbox/tex/jshort,/Users/seananderson/Dropbox/tex/ref3}

\clearpage

\section{Tables}

\input{sparks.tex}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{t-nu-eg.pdf}
\caption{
An illustration of fitting population dynamics models that allow for heavy tails, represented by the Student-t degrees of freedom parameter $\nu$.
(a, b) The probability density for t distributions with a scale parameter of 1 and different values of $\nu$.
Small values of $\nu$ create heavy tails.
As $\nu$ approaches infinity the distribution approaches the normal distribution.
For example, at $\nu = 2$, the probability of drawing a value less than -5 is 1.8\%, whereas the probability of drawing such a value from a normal distribution is nearly zero ($2.9\cdot10^{-5}$\%).
(c--e) Simulated population dynamics from a Gompertz model with process noise drawn from t distributions with different values of $\nu$.
Coloured dots in panels c and d represent jumps with less than a 1 in 1000 chance of occurring in a normal distribution.
(f--h) Estimates of $\nu$ from models fit to the times series in panels c--e.
Shown are the posterior samples (histograms), median and interquartile range of the posterior (IQR) (dots and line segments), and the exponential prior on $\nu$ (dashed lines).
Colour shading behind panels f--h illustrates the region of heavy tails.}
\label{fig:didactic}
\end{center}
\end{figure}

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.44\textwidth]{nu-coefs-2.pdf}
\caption{
Estimates of how heavy-tailed population dynamics are for \nuCoefPopN\ populations of birds, mammals, insects, and fishes.
Small values of $\nu$ ($\lesssim 10$) suggest heavy-tailed dynamics; larger values of $\nu$ suggest approximately normal-tailed dynamics.
Vertical points and line segments represent posterior medians and 50\% / 90\% credible intervals for individual populations.
Inset plots show probability that $\nu < 10$ (probability of heavy tails) for populations arranged by taxonomic order.
Taxonomic orders are sorted by decreasing mean p($\nu < 10$).
Taxonomic orders with three or fewer populations in panel a are omitted for space.
Red and orange colours highlight populations with a high probability of heavy-tailed dynamics.
}
\label{fig:nu-coefs}
\end{center}
\end{figure}

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{correlates-p10.pdf}
\caption{
Potential correlates of heavy-tailed behaviour (indicated by a high probability that $\nu < 10$).
Shown are (a--c) parameters from the Gompertz heavy-tailed population model, (b) number of time steps, (c) body length, and (d) lifespan.
For the Gompertz parameters, $\sigma$ refers to the scale parameter of the Student-t process-noise distribution, $\lambda$ refers to the population growth rate at an abundance of one, $b$ refers to the density dependence parameter ($1$ is maximally density independent, $0$ is maximally density dependent, and $<0$ is inversely density dependent).
%For parameter values, points and lines represent medians and 50\% credible intervals of the posterior.
Circles representing a limited number of sharks, crustaceans, and gastropods are filled in white.
Median and 90\% credible interval posterior predictions of a beta regression multilevel model are shown in panels a and d where there was a high probability the slope coefficient was different from zero (Supporting Material).
}
\label{fig:correlates}
\end{center}
\end{figure}