### 2016-05-26 text/ms.tex
Revise main text including adding new introduction paragraph

\section{Black-swan events in animal populations}

Sean C. Anderson\textsuperscript{1,2*}, Trevor A.
Branch\textsuperscript{3}, Andrew B. Cooper\textsuperscript{2}, Nicholas K.
Dulvy\textsuperscript{1}

\textsuperscript{1}Earth to Ocean Research Group, Department of Biological
Sciences, Simon Fraser University, Burnaby BC, V5A 1S6, Canada

\textsuperscript{2}School of Resource and Environmental Management, Simon
Fraser University, Burnaby, BC, V5A 1S6, Canada

\textsuperscript{3}School of Aquatic and Fishery Sciences, Box 355020,
University of Washington, Seattle, WA 98195, USA

\textsuperscript{*}Corresponding author: Sean C. Anderson; E-mail:
sandrsn@uw.edu; Present address: School of Aquatic and Fishery Sciences, Box
355020, University of Washington, Seattle, WA 98195, USA

Short title: Black-swan events in animal populations

Keywords:
Bayesian,
black swans,
die-offs,
ecological surprises,
heavy tails,
Gompertz,
mass mortality,
population dynamics,
risk,
Student t

Words in abstract: TOdO

Words in main text: todo

Number of references: todo

Number of figures: 4

Number of tables: 0

\newpage
%125 now is 150

\section{Abstract}

Black swans are improbable events that nonetheless
occur---often with profound consequences. While physical extremes, such as
monsoons and heat waves, are widely studied and increasing in magnitude and
frequency, it remains unclear the extent to which ecological population numbers
buffer or suffer from such extremes. Here we estimate the frequency and
direction of black-swan events in \NPops\ animal populations, finding strong
evidence for their occurrence in \(\sim\)\overallBasePerc \% of populations.
Black-swan events occur most frequently for birds (\birdPH \%), mammals
(\mammalsPH \%), and insects (\insectsPH \%), and manifest primarily as
population die-offs and crashes (\percBSDown \%) driven by extrinsic
perturbations from climate variability. Thus, ignoring black-swan events will
underestimate extinction risk. Our results demonstrate the importance of both
modeling heavy-tailed downward events in populations, and developing
conservation strategies that are robust to future increases in climate
extremes.

% \textbf{One Sentence Summary}: Black swans---statistically improbable events
% with profound consequences---occur in animal populations and understanding
% them will improve climate risk assessments.

%\clearpage

\section{Introduction}

Major surprises happen more often than expected in
financial, social, and natural systems \citep{taleb2007, sornette2009, may2008}.
Massive unpredictable market swings are responsible for the majority of
financial gains and losses \citep{taleb2007}, fatalities from the largest wars
dwarf those from all others  \citep{newman2005}, and the frequency of the most
damaging earthquakes has exceeded past expectations \citep{sornette2009}. In
ecological systems, background rates of global extinction are punctuated by
mass extinction \citep{harnik2012}, viruses can mutate suddenly to infect new
hosts, and billions of animals can die at once in mass mortality
events \citep{fey2015}.

Indeed, such die-offs may be the most important element
affecting population persistence \citep{mangel1994} and their importance is
likely to increase given projected increases in the frequency and magnitude of
climate-related extremes \citep{ipcc2012}.
% Despite this anecdotal evidence for
% ecological black swans, systematic searches of large numbers of timeseries have
% yet to reveal black-swan events \citep{keitt1998, allen2001, halley2002}.
However, the overwhelming majority of population model fitting and risk forecasting
assumes that deviations from model predictions can be represented by a normal
distribution \citep[typically on a log scale; e.g.][]{brook2006a, knape2012}.
If black swans occur, though, a normal distribution would under-estimate the
probability of extreme events occurring.

There are many examples of black-swan events in populations
but we lack a large-scale comparative analysis
with a method that allows for
population dynamics
(e.g.\ per capita productivity, density dependence, and typical stochasticity).
Previous comparative analyses fitting
heavy-tailed distributions to timeseries
(distributions with higher probabilities of extreme events
than the normal distribution)
have not accounted for an underlying population dynamics model
\citep{keitt1998, allen2001, halley2002, segura2013}.
Alternatively, most examples of population die offs come from identifying
jumps in abundance that exceed a threshold of decline
\citep[e.g.][]{young1994, gerber2001, reed2003, fey2015}.
But this approach gives no indication of relative frequency
and again does not distinguish events from
expected dynamics.
Methodologically, the generalized extreme value (GEV) distribution represents
the limit distribution of a series of maxima or minima.
Although the GEV is well-suited for physical ecological variables
\citep[e.g.\ sedimentation rates or wind speeds;][]{gaines1993, denny2009},
because the GEV requires maxima per time block
or peaks beyond a chosen threshold,
the GEV requires longer timeseries then typically available
for populations and is not simply integrated
into a population dynamics model.

Here, we develop a new approach to estimate the frequency and magnitude of
black-swan dynamics across time series of 609 populations from a wide array of
taxonomic groups---including many  birds, mammals, insects, and fishes
(Table~S1). We identify characteristics of time series or
life-history traits associated with the detection of black-swan events and
verify known causes. To accomplish this, we develop a framework for identifying
heavy-tailed (black-swan) process noise in population dynamics, i.e.~whether
the largest stochastic jumps in log abundance from one time step to the next
are more extreme than typically seen with a normal distribution. Our framework
allows for a range of population dynamic models, can incorporate observation
uncertainty and skewness in process noise, and can be readily applied to
abundance time series.

\section{Methods}

\subsection{Data}

We selected abundance time series from the Global Population Dynamics
Database \citep{gpdd2010} (GPDD), which contains nearly 5,000 time series of
abundance from \(\sim\) 1,000 species and \(\sim\) 100 taxonomic orders. We
filtered the data (Supporting Information) to remove populations from less
reliable data sources, and those without sufficient data for our models, and
then interpolated some missing values (sensu Brook \textit{et
al.} \citep{brook2006a}). Our interpolation affected only \(\sim\)
\interpPointsPerc \% of the final data points (Table~S1) and none of the data
points that were later considered black-swan events. Our final dataset
contained \NPops\ populations across \NOrders\ taxonomic orders and seven
taxonomic classes, with a median of \medianTimeSteps\ time steps (range
\minTimeSteps--\maxTimeSteps) (Table~S1).

\subsection{Modeling framework}

We fit heavy-tailed Gompertz population dynamics models to data from the GPDD. The Gompertz model represents population growth as a linear function in log space. If \(x_{t}\) represents the ln abundance (\(N\)) at time \emph{t} then
\begin{align}
x_t &= \lambda + bx_{t - 1} + \epsilon_t\\
\epsilon_t &\sim \mathrm{Student\mhyphen t}(\nu, 0, \sigma)
\end{align}
The growth parameter \(\lambda\) represents the expected growth rate if \(N_{t}
= 1\). The model is density independent if \(b = 1\), maximally density
dependent if \(b = 0\), and inversely density dependent if \(b < 0\). Usually,
the process noise \(\epsilon_{t}\) is modelled as normally distributed, but we
allow it to be drawn from a Student-t distribution with scale parameter
\(\sigma\) and degrees of freedom \(\nu\). We can then estimate the degree to
which the process deviations have heavy tails and are therefore evidence of
black-swan events (Fig.~1A, B).

For the Gompertz model, our weakly-informative priors (Fig.~\ref{fig:priors}) were:
\begin{align}
b &\sim \mathrm{Uniform}(-1, 2)\\ \lambda &\sim \mathrm{Normal}(0, 10^2)\\
\sigma &\sim \mathrm{Half\mhyphen Cauchy} (0, 2.5)\\ \nu &\sim
\mathrm{Truncated\mhyphen Exponential}(0.01, \mathrm{min.} = 2).
\end{align}
Our prior on \(b\) was uninformative between values of \(-1\) and \(2\). We
would not expect values of \(b\) with levels of density dependence as low as
\(-1\) (very strong inverse density dependence), nor would we generally expect
values above \(1\). We allowed values of \(b\) above \(1\) to allow for
non-stationary time series of growth rates. The estimates of \(b\) were well
within these bounds. Our prior on \(\lambda\) was weakly informative within the
range of expected values for population growth and is similar to the default
priors suggested by Gelman \emph{et al.} \citep{gelman2008d} for intercepts of
regression models. Our Half-Cauchy prior on \(\sigma\) follows Gelman \emph{et
al.} \citep{gelman2006c} and Gelman \emph{et al.} \citep{gelman2008d} and the
specific scale parameter of \(2.5\) is based on our expected range of the value
in nature from previous studies \citep{connors2014}. In our testing of a
subsample of populations, our parameter estimates were not qualitatively
changed by switching to a uniform prior on \(\sigma\), but the weakly
informative Half-Cauchy prior sped up chain convergence.

We chose a weakly informative exponential prior on \(\nu\) with rate parameter
of \(0.01\) truncated at values above two---a slightly less informative prior
than suggested by Fernandez and Steel \citep{fernandez1998}. We truncated the
distribution at \(2\), since at \(\nu < 2\) the variance of the t distribution
is undefined. This prior gives only a \(7.7\)\% probability that \(\nu < 10\)
but constrains the sampling sufficiently to avoid drifting towards infinity. In
any case, for \(\nu > 20\) the t distribution is almost indistinguishable from
the normal distribution (Fig.~1). Based on the shape of the t distribution, we
chose the probability that \(\nu < 10\), Pr(\(\nu < 10\)), to define the
probability of heavy-tailed dynamics. When categorizing a population as heavy
tailed, we applied a 0.5 probability threshold. In the scenario where the data
are uninformative about heavy tails (e.g.~Fig.~1E, H), the posterior for
\(\nu\) will approximately match the prior (prior median \(= 71\), mean \(=
102\)) and a metric of Pr\((\nu < 10) > 0.5\) is unlikely to flag the
population as heavy tailed.

We fit our models with Stan 2.6.0 \citep{stan-manual2015, hoffman2014}, and R
3.2.0 \citep{r2015}. We began with four chains and 2,000 iterations, discarding
the first 1,000 as warm up (i.e.~4,000 final samples). If \(\hat{R}\) (the
potential scale reduction factor---a measure of chain convergence) was greater
than \(1.05\) for any parameter, or if the minimum effective sample size,
\(n_\mathrm{eff}\) (a measure of the effective number of uncorrelated samples),
for any parameter was less than \(200\), we doubled both the total iterations
and warm up period and sampled from the model again. These thresholds are in
excess of the minimums recommended by Gelman \emph{et al.} \citep{gelman2006a} of
\(\hat{R} < 1.1\) and \(n_\mathrm{eff} > 100\) for reliable point estimates and
confidence intervals. In the majority of cases our minimum thresholds were
greatly exceeded. We continued this procedure up to 8,000 iterations (16,000
total samples) by which all chains were deemed to have sufficiently converged.
The No-U-Turn Hamiltonian Markov chain Monte Carlo sampler in Stan generally
requires fewer iterations to obtain equivalent effective sample sizes than the
Gibbs or Metropolis--Hastings algorithms \citep{stan-manual2015}.

Specifically, we fit population dynamics models in which the process noise is
drawn from a Student-t distribution. By estimating the degrees of freedom
parameter, \(\nu\), we can estimate the degree to which the process deviations
have heavy tails and are therefore diagnostic of black-swan events (Fig.~1,
Fig.~S\ref{fig:priors}). Lower values of \(\nu\) result in heavy-tailed
distributions. For example, at \(\nu = 2\), the probability of drawing a value
more than five standard deviations below the mean is almost 70,000 times higher
than the probability of drawing such a value from a normal distribution
(\(0.02\) vs.\ \(2.9\cdot10^{-7}\)). As \(\nu\) approaches infinity, the
distribution approaches the normal distribution (Fig.~1).

\subsection{Alternative priors}

To test if the prior on \(\nu\) influenced our estimate of black-swan dynamics,
we refit our models with weaker and stronger priors. Our base model used a
prior on \(\nu\) of Truncated-Exponential(0.01, min.\ = 2). For a weaker prior
we used Truncated-Exponential(0.005, min.\ = 2) and for a stronger prior we
used Truncated-Exponential(0.02, min.\ = 2) (Fig.~\ref{fig:priors}c). Note that
the base and weaker priors are relatively flat within the region of \(\nu <
20\), which is the region we are mostly concerned about when categorizing
populations as heavy tailed. We furthermore considered a Uniform(\(0, 1\))
prior on \(1/\nu\) as used in Gelman \emph{et al.} \citep{gelman2014}, but this
prior puts considerably more weight on lower (heavy-tailed) values of \(\nu\)
than the truncated exponential priors considered here (not shown).

\subsection{Alternative population models} We fit alternative population models to
test if four key phenomena systematically changed our conclusions. The range of
percentages of black swans by taxonomic class cited in the abstract are based
on lower and upper limits across our main Gompertz model and these four
alternative models.

\subsubsection{Autocorrelated residuals} We considered a version of the Gompertz
model in which an autoregressive parameter was fit to the process-noise
residuals: \begin{align} x_t &= \lambda + b x_{t-1} + \epsilon_t\\ \epsilon_t
&\sim \mathrm{Student\mhyphen t}(\nu, \phi \epsilon_{t-1}, \sigma). \end{align}
In addition to the parameters in the original Gompertz model, this model
estimates an additional parameter \(\phi\), which represents the correlation of
subsequent residuals. Based on the results of previous analyses with the
GPDD \citep{connors2014} and the chosen prior in previous
analyses \citep{thorson2014a}, we placed a moderately informative prior on
\(\phi\) that assumed the greatest probability density near zero with the
reduced possibility of \(\phi\) being near \(-1\) or \(1\). Specifically, we
chose \(\phi \sim \mathrm{Truncated\mhyphen Normal}(0, 0.5, \mathrm{min.} = -1,
\mathrm{max.} = 1)\).

\subsubsection{Density independence} We fit a simplified version of the
Gompertz model in which the density dependence parameter \(b\) was fixed at
\(1\) (density independent). This is equivalent to fitting a random walk model
(with drift) to the \(\ln\) abundances or assuming the growth rates are drawn
from a stationary distribution. The model was as follows:
\begin{align}
x_t &= \lambda + x_{t-1} + \epsilon_t\\
\epsilon &\sim \mathrm{Student\mhyphen t}(\nu, 0, \sigma).
\end{align}
We fit this model for three reasons: (1) it is computationally simpler and so
provides a check that our more complicated full Gompertz model was obtaining
reasonable estimates of \(\nu\), (2) it provides a test of whether density
dependence was systematically affecting our perception of heavy tails, (3) it
matches how some previous authors have modelled heavy tails without accounting
for density dependence \citep{segura2013}.

\subsubsection{Observation error} Observation error can bias parameter
estimates \citep{knape2012} and is known to affect the ability to detect extreme
events \citep{ward2007}. In our main analysis, we fit a model that ignored
observation error. One way to account for observation error would be to fit a
full state-space model that simultaneously estimates the magnitude of process
noise and observation error. However, simultaneously estimating observation and
process noise is a challenging problem (e.g.\ because the observation and
process noise parameters tend to negatively covary in model fitting) and is
known to sometimes result in identifiability issues with the Gompertz
population model \citep{knape2008}. Furthermore, our model was applied to
hundreds of time series, often of short length (as few as \minTimeSteps\ time
steps) and our model estimates an additional parameter---the shape of the
process deviation tails---potentially making identifiability and computational
issues even greater. Therefore, we considered a version of the base Gompertz
model that allowed for a fixed level of observation error:
\begin{align}
U_t &= \lambda + b U_{t-1} + \epsilon_t\\
x_t &\sim \mathrm{Normal}(U_t, \sigma_\mathrm{obs}^2)\\
\epsilon_t &\sim \mathrm{Student\mhyphen t}(\nu, 0, \sigma_\mathrm{proc}),
\end{align}
where \(U\) represents the unobserved state vector, \(\sigma_\mathrm{obs}\)
represents the standard deviation of observation error (on a log scale), and
\(\sigma_\mathrm{proc}\) represent the process noise scale parameter. We set
\(\sigma_\mathrm{obs}\) to \(0.2\), which represents a typical value used in
simulation analyses \citep{valpine2002, thorson2014b}.

\subsubsection{Ricker-logistic} We also fit a Ricker-logistic model:
\begin{align}
x_t &= x_{t-1} + r_{\mathrm{max}}\left(1 - \frac{N_{t-1}}{K}\right) + \epsilon_t\\
\epsilon_t &\sim \mathrm{Student\mhyphen t}(\nu, 0, \sigma),
\end{align}
where \(r_\mathrm{max}\) represents the theoretical maximum population growth
rate that is obtained when \(N_t = 0\). The parameter \(K\) represents the
carrying capacity and, as before, \(x_t\) represents the \(\ln\) transformed
abundance at time \(t\). The Ricker-logistic model assumes a linear decrease in
population growth rate with increases in abundance. In contrast, the Gompertz
model assumes a linear decrease in population growth rate with increases in
\textit{log} abundance.

To fit the Ricker-logistic models, we chose a prior on \(K\) uniform between
zero and twice the maximum observed abundance (similar and less informative
than in \citet{delean2013}). We set the prior on
\(r_\mathrm{max}\) as uniform between 0 and 20 as in \citet{delean2013}. We used the same priors on \(\nu\) and \(\sigma\) as in
the Gompertz model.

\subsection{Simulation testing}
We performed two types of simulation testing.
First, we tested how easily the Student-t distribution \(\nu\) parameter could
be recovered given different true values of \(\nu\) and different sample sizes.
Second, we tested the ability of the heavy-tailed Gompertz model to obtain
unbiased parameter estimates of \(\nu\) given that a set of process deviations
was provided in which the effective \(\nu\) value was close to the true \(\nu\)
value.

We found that we could consistently and precisely recover median posterior
estimates of \(\nu\) near the true value of \(\nu\) with large samples (\(\ge
200\)) (Fig.~\ref{fig:sim-nu} upper panels). At smaller samples we could
usually qualitatively distinguish heavy from not-heavy tails, but the model
tended to underestimate how heavy the tails were. At the same time, at smaller
sample sizes, the model tended to overestimate how large the scale parameter
was (Fig.~\ref{fig:sim-nu} lower panels).

Our simulation-testing results indicate that the Gompertz model can recapture
the true value of \(\nu\) when the process noise was chosen so that appropriate
tail events were present (Fig.~\ref{fig:sim-prob} upper panels). The addition
of observation error caused the model to tend to underestimate the degree of
heavy-tailedness. Fitting a model with correctly specified observation error
did not make substantial improvements to model bias (Fig.~\ref{fig:sim-prob}).

\subsection{Covariates of heavy-tailed dynamics} We fit a hierarchical beta
regression model to the predicted probability of heavy tails, Pr(\(\nu < 10\)),
to investigate potential covariates of heavy-tailed dynamics. We obtained
maximum lifespan and body-size data from \citet{brook2006a}.
The beta distribution is useful when response data are on a continuous scale
between zero and one \citep{ferrari2004}. The model was as follows:
\begin{align}
\mathrm{Pr}(\nu_i < 10) &\sim \mathrm{Beta}(A_i, B_i)\\
\mu_i &= \mathrm{logit}^{-1}(\alpha
  + \alpha^\mathrm{class}_{j[i]}
  + \alpha^\mathrm{order}_{k[i]}
  + \alpha^\mathrm{species}_{l[i]}
  + X_i \beta),
  \: \text{for } i = 1, \dots, 617\\
A_i &= \phi_\mathrm{disp} \mu_i\\
B_i &= \phi_\mathrm{disp} (1 - \mu_i)\\
\alpha^\mathrm{class}_j &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{class}}),
  \: \text{for } j = 1, \dots, 6\\
\alpha^\mathrm{order}_k &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{order}}),
  \: \text{for } k = 1, \dots, 38\\
\alpha^\mathrm{species}_l &\sim
  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{species}}),
  \: \text{for } l = 1, \dots, 301,
\end{align}
where \(A\) and \(B\) represent the beta distribution shape parameters;
\(\mu_i\) represents the predicted value for population \(i\), class \(j\),
order \(k\), and species \(l\); \(\phi_\mathrm{disp}\) represents the
dispersion parameter; and \(X_i\) represents a vector of predictors (such as
lifespan) for population \(i\) with associated \(\beta\) coefficients. The
intercepts are allowed to vary from the overall intercept \(\alpha\) by
taxonomic class (\(\alpha^\mathrm{class}_j\)), taxonomic order
(\(\alpha^\mathrm{order}_k\)), and species (\(\alpha^\mathrm{species}_l\)) with
standard deviations \(\sigma_{\alpha \; \mathrm{class}}\), \(\sigma_{\alpha \;
\mathrm{order}}\), and \(\sigma_{\alpha \; \mathrm{species}}\). Where possible,
we also allowed for error distributions around the predictors by incorporating
the standard deviation of the posterior samples for the Gompertz parameters
\(\lambda\), \(b\), and \(\log \sigma\) around the mean point value as normal
distributions (not shown in the above equation).

We log transformed \(\sigma\), time-series length, and lifespan to match the
way they are visually represented in Fig.~\ref{fig:correlates} and to make the
relationship approximately linear on the logit-transformed response scale. All
input variables were standardized by subtracting their mean and dividing by two
standard deviations to make their coefficients comparable in
magnitude \citep{gelman2008c}. We excluded body length as a covariate because it
was highly correlated with lifespan, and lifespan exhibited more overlap across
taxonomy than body length. Lifespan is also more directly related to time and
potential mechanisms driving black-swan dynamics.

We incorporated weakly informative priors into our model: \(\mathrm{Cauchy}(0,
10)\) on the global intercept \(\alpha\), \(\mathrm{Half\mhyphen Cauchy}(0,
2.5)\) on all standard deviation parameters, \(\mathrm{Half\mhyphen Cauchy}(0,
10)\) on the dispersion parameter \(\phi_\mathrm{disp}\), and
\(\mathrm{Cauchy}(0, 2.5)\) on all other parameters \citep{gelman2006c,
gelman2008d}. Compared to normal priors, the Cauchy priors concentrate more
probability density around expected parameter values while allowing for a
higher probability density far into the tails, thereby allowing the data to
dominate the posterior more strongly if it disagrees with the prior. We fit our
models with 5,000 total iterations per chain, 2500 warm-up iterations, four
chains, and discarding every second sample. We checked for chain convergence
visually and with the same criteria as before (\(\widehat{R} < 1.05\) and
\(n_\mathrm{eff} >200\) for all parameters). We further checked our model
against a similar model with the glmmADMB \textsf{R} package \citep{glmmadmb}
without uncertainty around the predictors.

To derive taxonomic-order-level estimates of the probability of heavy tails
accounting for time-series length (Fig.~3A), we fit a separate hierarchical
model with the same structure but with only \(\log\) time-series length as a
predictor---in this case, we did not want to control for intrinsic population
characteristics such as density dependence. Since our predictors were centered
by subtracting their mean value, we obtained order-level estimates of the
probability of heavy tails at mean log time-series length by adding the
posteriors for \(\alpha\), \(\alpha^\mathrm{class}_j\), and
\(\alpha^\mathrm{order}_k\).

\subsection{Skewed Student-t forecasts} To formally evaluate the apparent skewness
of heavy-tailed process noise we fit Gompertz models with skewed Student-t
distributed process noise \citep{fernandez1998}. The distribution adds one
parameter to the Student-t distribution, \(\gamma\), which controls the
skewness. The distribution is symmetrical if \(\gamma = 1\), left skewed if \(0
< \gamma < 1\), and right skewed if \(1 < \gamma < \infty\). We placed a weakly
informative prior of \(\mathrm{Cauchy}(0, 2.5)\) on \(\log \gamma\). We
aggregated 5,000 randomly selected posterior samples from the \(\gamma\)
parameter of each model at three levels of evidence for heavy-tails:
\emph{heavy tailed} median \(\nu < 10\); \emph{slightly heavy tailed} \(10 \le
\mathrm{median}\, \nu \le 70\), and \emph{normal tails}, median \(\nu > 70\)
(Fig.~4).

To generate five-year forecasts of abundance, we combined the posterior
parameter samples from the skewed-t Gompertz models with stochastically
generated process noise. We compared these forecasts to those generated by a
standard Gompertz model with normally distributed process noise fit to the same
data. We calculated the ratio of the lower 99\% quantile credible interval
between the two model projections for all populations in which Pr(\(\nu < 10\))
\(> 0.5\). To ensure the results had stabilized in the tails of the forecast
posterior, we increased the number of posterior samples in Stan. We ran four
chains of 20,000 iterations and discarded the first 10,000 as warmup for a
total of 40,000 samples.

\section{Results}

We found evidence that black-swan dynamics are infrequent but highly
influential: they occur most frequently for birds (7\%), mammals (5\%), and
insects (3\%) and almost never in fishes for this dataset (Fig.~2). Black-swan
events were taxonomically widespread, occurring in 38\% of taxonomic orders.
Accounting for time series length and partially pooling inference across
taxonomic class and order with a hierarchical model, we found stronger evidence
for black swans in insect populations than these statistics suggest---four of
eight orders with the highest median probability of heavy tails were insect
orders (Fig.~3a).

The majority of our heavy-tailed estimates were robust to alternative
population models, observation error, and choice of Bayesian priors. Our
conclusions were not systematically altered when we included an autocorrelation
structure in the residuals, modelled population growth rates with or without
density dependence, or modelled the population dynamics as Ricker-logistic
instead of Gompertz (Fig.~S\ref{fig:alt}). Similarly, introducing moderate
observation error (CV = 0.2) only slightly decreased the estimated prevalence
of black-swan events (Fig.~S\ref{fig:alt}), and the strength of the prior on
\(\nu\) had little influence on estimates of black-swan dynamics
(Fig.~S\ref{fig:alt-priors}). Finally, our simulation testing shows that, if
anything, our models underpredict the true magnitude and probability of
heavy-tailed events---especially given the time series have a median of only 26
years in our analysis (Figs~S\ref{fig:sim-nu},~S\ref{fig:sim-prob}).

% Our results show that weaker and stronger priors have little influence on our
% conclusions about heavy-tailed dynamics (Fig.~\ref{fig:alt-priors}). When the
% data are informative about tail behaviour (i.e. when there is strong evidence
% of low \(\nu\) values, upper-right of Fig.~\ref{fig:alt-priors}), the prior has
% little impact on the estimate of \(\nu\). When the data are less informative
% about \(\nu\) (i.e.\ when there are no or few tail events and time series are
% short or noisy), the prior can pull the estimate of \(\nu\) towards larger or
% smaller values (Fig.~\ref{fig:alt-priors}). The vast majority of the
% populations with Pr\((\nu < 10)\) in the base prior were not altered
% qualitatively by this range of prior strength.

For model fits to the population data, the probability of detecting black-swan
dynamics was positively related to time-series length and negatively related to
magnitude of process noise but not clearly related to population growth rate,
density dependence, or maximum lifespan (Fig.~3b, Fig.~S\ref{fig:correlates}).
Longer time-series length was the strongest covariate of observing black-swan
dynamics: there is a 1.6 times greater probability of detecting a black-swan
event in populations with 60 time steps than in one with 30 time steps
(Fig.~S\ref{fig:correlates}).

The majority of black-swan events (\percBSDown \%) were downward (die-offs)
rather than upward (unexpectedly rapid abundance increases). Of the black-swan
events with published explanations (Table~S2), the majority involved
a combination of multiple factors. For example, a synchronization of
environmental- and predation-mediated population cycles is thought to have
caused a downward black-swan event for a water vole (\emph{Arvicola
terrestris}) population \citep{saucy1994}. Other black swans were the result of
a sequence of extreme climate events on their own. For instance, severe winters
in 1929, 1940--1942, and 1962--1963 were associated with black-swan downswings
in grey heron (\emph{Ardea cinerea}) abundance in the United
Kingdom \citep{stafford1971} (Fig.~4c). Our analysis finds that the last event
was a combination of two heavy-tailed events in a row and the population took
three times longer to recover than predicted \citep{stafford1971}. Downward black
swans were sometimes followed by upward black swans. For example, during
a period of population crowding and nest shortages, a population of European
shag cormorants (\emph{Phalacrocorax aristotelis}) on the Farne Islands, United
Kingdom declined suddenly following a red tide event in 1968 \citep{potts1980}.
This freed up quality nest sites for first-time breeders, productivity
increased, and the population experienced a rapid upswing in
abundance \citep{potts1980}.

Given the prevalence of downwards events, we refit our heavy-tailed models to
measure the degree of skewness \(\gamma\) of the process noise using a skew-t
distribution (Fig.~4a), and used these models to make near-term risk forecasts.
Aggregated across populations with strong evidence of heavy tails (median \(\nu
< 10\)), 85\% of the \(\gamma\) probability density was below 1, indicating
strong evidence of downwardly skewed process noise (Fig.~4b,
Fig.~S\ref{fig:skew-nu}). In contrast, populations that did not have heavy
tails (median \(\nu \geq 70\)) had little evidence of skewed process noise
(Fig.~4b; 90\% of 95\% credible intervals overlapped 1). Projecting these
heavy-tailed populations forward five years revealed that assuming standard
normal process noise underestimated risk (99\% lower credible interval of
abundance) by \crashUnderRange-fold (interquartile range; Fig.~4c,
Fig.~S\ref{fig:skew-projections}).

% todo Thus our analysis provides strong evidence
% for downward-skewed heavy-tailed events in abundance time series of higher
% taxa, and ignoring these events will tend to underestimate the risk of
% population declines.

\section{Discussion}

There are a number of possible causes of black-swan events, including
unmodelled intrinsic properties of populations or extrinsic forces acting on
populations. For example, we could observe black-swan dynamics if we miss an
underlying mixture of processes, since a mixture of normal distributions with
different variances can generate a t distribution with heavy
tails \citep{allen2001}. In ecology, these processes could arise through an
aggregation of populations across space, population
diversity \citep{schindler2010}, or an intrinsic change in population variability
through time \citep{carpenter2006}. Extrinsic ecological forces could also cause
black-swan dynamics \citep{nunez2012} including extreme climate
events \citep{meehl2004, katz2005, ipcc2012}, predation from (or competition
with) other species experiencing black swans, or sharp changes in human
pressures such hunting, fishing, or habitat destruction. Alternatively, the
synchrony of multiple extrinsic forces could give rise to black-swan dynamics
through synergistic interactions \citep{kirby2009} or a rare alignment of
non-synergistic forces \citep{denny2009}.

Given the multitude of reasons why a population modeler might expect heavy tail process noise, what should a population modeler do?

Essentially any reason for why a population modeler has incorrectly specified
the exact process generating the abundance times there is under question could
generate have detailed process noise.

We think modelers should routinely consider substituting a heavy tailed
distribution, such as the student t-distribution, to represent the process
noise in population dynamic modeling. Especially when making forecasts to make
predictions about risk of specified bad events.

But what heavy tailed distribution to pick?

This is an open research question. However, we think the student t-distribution
represents a simple solution in that it can be easily substituted for the
normal distribution and if the degrees of freedom are fixed at a moderately
heavy level it doesn't require the estimation of any extra parameters over the
normal distribution.

And if we use a student t-distribution than what degrees of freedom parameter should be you assume?

The exact shape of the tails has been shown to be important in the field of
dispersal biology -- has a big effect on dynamic \citep{kot1996}
\citep{clark1999}
\citep{kot1996}

The degrees of
freedom could be estimated from meta-analysis but even here using one of the
best large data sets of animal abundance timeseries the timeseries are
relatively short and the taxonomy is a biased sample of species of interest to humans  economic
interest ...

for example the majority of the blocks one offense that we observed were
population crashes This could be due to basic biology --- animal birthrates are
constrained whereas every animal can die --- on the other hand it could be due
to the biased sample of species in the data set which represent species of
concern to humans and we don't have much data for highly reproductive species
that are not threatened or of economic interest

If there is any reason for a model are to expect that external forces could act
or that there is a mixture of distributions through spatial heterogeneity we
think it is prudent to err on the side of caution and allow for heavy tailed
process noise in forecasts

The barring a better estimate we might suggest using a degrees of freedom parameter around 5

this allows for a catastrophic event of an x fold decline of abundance that
happens every Y years compared to a normal distribution allowing only for this
event to happen every Z years.

But what was once considered a one in X million event then becomes a one in Y
million event

A t-distribution compared to a normal distribution on the process error he
essentially lets us be right most of the time while allowing are models to be
very wrong occasionally this better matches our current intuition that
ecological surprises can and do occur

Our hypothesis is that our first priority is to allow for tails that are
heavier than the normal distribution and approximately in line with those
discovered in this paper and our second priority is to determine exactly how
heavy a default distribution we should assume.

Answers might come from simulation analyses exploring the implications of
different tail shapes and also from fitting population models to known time
series of mass mortality events \citep[e.g.][]{fey2015}

In light of our findings, we suggest that ecological resource management can
learn from disciplines that focus on heavy tails. For example, earthquake
preparedness and response is focused on black-swan events. Similarly to
ecological black swans, we can rarely predict the specific timing of large
earthquakes. But, earthquake preparedness involves spatial planning based on
forecast probabilities to focus early detection efforts and develop disaster
response plans \citep{nrc2007}. The presence of black swans also suggests that we
develop management policy that is robust to heavy tails and encourages
\textit{general} resilience \citep{carpenter2012}. For instance, setting target
population abundances far from critical limits will buffer against black-swan
events \citep{caddy1996}, and maintaining genetic, phenotypic, and behavioural
diversity may allow some components of populations to persist when others are
affected by disease or extreme environmental forces \citep{schindler2010}.
Finally, surprising, or counterintuitive ecological dynamics offer a tremendous
opportunity to learn about ecological systems, evaluate when models break down,
and adjust future management policy \citep{doak2008, lindenmayer2010}.

Rare catastrophes can have a profound influence on population
persistence \citep{mangel1994}. In recent decades, ecology has moved toward
focusing on aspects of variance in addition to mean
responses \citep{thompson2013}. Our results suggest that an added focus on
ecological extremes represents the next frontier, particularly in the face of
increased climate extremes \citep{meehl2004, ipcc2012, thompson2013}. Financial
analysts are concerned with the shape of financial return downward tails
because these directly impact estimates of risk---the probability of a specific
magnitude of undesired event occurring. Similarly, ecologists should focus more
on estimating and predicting downward tails of population abundance, since
these increase true extinction risk.

\section{Acknowledgements}

We thank J.W. Moore, A.O. Mooers, L.R. Gerber, J.D.
Yeakel, C. Minto and members of the Earth to Ocean Group for helpful
discussions and comments. We are grateful to the contributors and maintainers
of the Global Population Dynamics Database and to Compute Canada's WestGrid
high-performance computing resources. Silhouette images were obtained from
phylopic.org under Creative Commons licenses; sources are listed in the
Supporting Information. Funding was provided by a Simon Fraser University
Graduate Fellowship (S.C.A.), the Natural Sciences and Engineering Research
Council of Canada (N.K.D., A.B.C.), the Canada Research Chairs Program
(N.K.D.).

\section{References}

\renewcommand{\section}[2]{}%

\bibliographystyle{ecologyletters}\bibliography{jshort,ms}

% \textbf{Supporting Information}\\
% Materials and Methods\\
% Supplementary References\\
% Tables S1 to S2\\
% Figs.~S1 to S8\\

%\textbf{Author Contributions}: S.C.A. and T.A.B. conceived the project;
%S.C.A., T.A.B., A.B.C., and N.K.D. designed the study; S.C.A. analyzed the
%data and wrote the paper with input from all authors.

\clearpage

% \section{Figures}

\begin{center}
\includegraphics[width=\textwidth]{../analysis/t-nu-eg2.pdf}
\end{center}

\textbf{Fig.~1: Illustration of population dynamic models that allow for heavy
tails.} (\textbf{A} to \textbf{B}) Probability density for the Student-t
distribution with scale parameter of 1 and different values of \(\nu\). Small
values of \(\nu\) create heavy tails while as \(\nu\) approaches infinity the
distribution approaches the normal distribution. (\textbf{C} to \textbf{E})
Simulated population dynamics from a Gompertz model with process noise drawn
from Student-t distributions with three values of \(\nu\). Coloured dots in
panels \textbf{C} and \textbf{D} represent jumps with less than a 1 in 1000
chance of occurring in a normal distribution. (\textbf{F} to \textbf{H})
Estimates of \(\nu\) from models fit to the times series in \textbf{C} to
\textbf{E}. Shown are posterior samples (histograms), median and interquartile
range of the posterior (IQR, dots and line segments), and the exponential prior
on \(\nu\) (dashed lines). Colour shading behind \textbf{F} to \textbf{H}
illustrates the approximate region of heavy tails.

\clearpage

\begin{center}
\includegraphics[width=0.36\textwidth]{../analysis/nu-coefs-2.pdf}
\end{center}

\textbf{Fig.~2: Estimates of population dynamic heavy-tailedness for
populations of birds, mammals, insects, and fishes.} (\textbf{A} to \textbf{D})
Small values of \(\nu\) suggest heavy-tailed black-swan dynamics. Vertical
points and line segments represent posterior medians and 50\% / 90\% credible
intervals for individual populations. Inset plots show probability that \(\nu
< 10\) for populations arranged by taxonomic order and sorted by decreasing
mean Pr(\(\nu < 10\)). Taxonomic orders with three or fewer populations in
\textbf{A} are omitted for space. Red to yellow points highlight populations
with a high to moderately high probability of heavy-tailed dynamics.

\clearpage

\begin{center}
\includegraphics[width=0.35\textwidth]{../analysis/order-posteriors-covariates.pdf}
\end{center}

\textbf{Fig.~3: Standardized probabilities and covariates of heavy-tailed
dynamics.} (\textbf{A}) Taxonomic-order-level posterior densities of Pr(\(\nu
< 10\)) after accounting for time-series length in a hierarchical model. Dotted
vertical line in a indicates the median expected Pr(\(\nu < 10\)) from the
prior distribution. Colour shading refers to taxonomic class (yellow: fishes,
green: insects, purple: birds, and red: mammals). Estimates are standardized at
the geometric mean of time series length across all the data (approximately 27
time steps) and shown for orders with \(\ge 5\) populations. (\textbf{B})
Posterior densities for potential covariates of Pr(\(\nu < 10\)). In both
panels, short vertical line segments within the density polygons indicate
medians.

\clearpage

\begin{center}
\includegraphics[width=0.65\textwidth]{../analysis/skew-fig.pdf}
\end{center}

\textbf{Fig.~4: Heavy-tailed process noise tends to be downward skewed and
ignoring this can underestimate risk.} (\textbf{A}) Illustrations of Student-t
distributions with 3 levels of skewness (\(\gamma\)) and heavy-tails (\(\nu\)).
(\textbf{B}) Posterior density of the skewness parameters aggregated across
populations grouped into heavy-tailed (\(\hat{\nu} < 10\)), slightly
heavy-tailed (\(10 \leq \hat{\nu} \leq 70\)), and normal-tailed populations
(\(\hat{\nu} > 70\)). Approximate mid-values from \textbf{B} are illustrated in
\textbf{A}. (\textbf{C}) Example time series of heavy-tailed populations with
different levels of skewness. Red dots highlight likely heavy-tailed events.
Forecasts (grey regions) show median (solid lines) and lower 99\% credible
intervals (dotted lines) of abundance. Black and red lines indicate forecasts
from Gompertz models with lognormal and skew-t process noise, respectively.