### 2017-02-17 analysis/phylopic.tex
Switch silhouettes

% latex table generated in R 3.3.1 by xtable 1.8-2 package
Taxonomic order & Credit & License URL \\
  \midrule
Gadiformes & Uncredited & \url{http://creativecommons.org/publicdomain/mark/1.0/} \\
  % Salmoniformes & Servien (vectorized by T. Michael Keesey) & \url{http://creativecommons.org/licenses/by-sa/3.0/} \\
  Perciformes & Ellen Edmonson and Hugh Chrisp (vectorized by T. Michael Keesey) & \url{http://creativecommons.org/publicdomain/mark/1.0/} \\
  Pleuronectiformes & Birgit Lang & \url{http://creativecommons.org/licenses/by/3.0/} \\
  Lepidoptera & Curtis (modified by T. Michael Keesey) & \url{http://creativecommons.org/publicdomain/mark/1.0/} \\
  Carnivora & Brian Gratwicke (photo) and T. Michael Keesey (vectorization) & \url{http://creativecommons.org/licenses/by/3.0/} \\
  Coleoptera & Crystal Maier & \url{http://creativecommons.org/licenses/by/3.0/} \\
  Rodentia & Mattia Menchetti & \url{http://creativecommons.org/publicdomain/zero/1.0/} \\
  Odonata & Gareth Monger & \url{http://creativecommons.org/licenses/by/3.0/} \\
  Lagomorpha & Sarah Werning & \url{http://creativecommons.org/licenses/by/3.0/} \\
  Anseriformes & Sharon Wegner-Larsen & \url{http://creativecommons.org/publicdomain/zero/1.0/} \\
  Passeriformes & Michael Scroggie & \url{http://creativecommons.org/publicdomain/zero/1.0/} \\
  Artiodactyla & Jan A. Venter, Herbert H. T. Prins, David A. Balfour and Rob Slotow (vectorized by T. Michael Keesey) & \url{http://creativecommons.org/licenses/by/3.0/} \\
  Diptera & Gareth Monger & \url{http://creativecommons.org/licenses/by/3.0/} \\
  Charadriiformes & Rebecca Groom & \url{http://creativecommons.org/licenses/by/3.0/} \\
  Hemiptera & T. Michael Keesey & \url{http://creativecommons.org/publicdomain/zero/1.0/} \\
  Falconiformes & Ferran Sayol & \url{http://creativecommons.org/publicdomain/zero/1.0/} \\
  Galliformes & Steven Traver & \url{http://creativecommons.org/publicdomain/zero/1.0/} \\
   \bottomrule

% Salmoniformes
% Lepidoptera http://phylopic.org/image/0fde2bb6-0472-4273-bf46-2d6073fa8fbc/
% Pleuronectiformes http://phylopic.org/image/4c2220b9-68a8-4594-8dd8-229643e11c84/

     % taxonomic_order                                                             url
%1         Gadiformes http://phylopic.org/image/8d92b454-3131-4bbd-ac9c-e1df14c2fc5a/
%3        Perciformes http://phylopic.org/image/426732d4-17c9-4e85-9672-b6b7f47eb99d/
%4       Anseriformes http://phylopic.org/image/cf522e02-35cc-44f5-841c-0e642987c2e4/
%5        Lepidoptera http://phylopic.org/image/0fde2bb6-0472-4273-bf46-2d6073fa8fbc/
%6          Carnivora http://phylopic.org/image/20da6c7c-2584-4cee-921b-ebd09384567b/
%7           Rodentia http://phylopic.org/image/bd3d2bc1-915c-4fa6-99b8-9a5fc6b725e0/
%8         Lagomorpha http://phylopic.org/image/dea688b6-9168-4e79-a106-366888148eb1/
%9         Coleoptera http://phylopic.org/image/77164e4b-8076-447b-a140-7a1f9f907da8/
%10           Odonata http://phylopic.org/image/ce57b692-e1e0-468e-a025-8c10ceebca22/
%11     Passeriformes http://phylopic.org/image/3d38baad-d355-432a-944c-07608828d0a4/
%12      Artiodactyla http://phylopic.org/image/bb357997-4d4d-4e73-b853-d95a7a47ec68/
%13 Pleuronectiformes http://phylopic.org/image/4c2220b9-68a8-4594-8dd8-229643e11c84/
%14     Falconiformes http://phylopic.org/image/f68e0e0e-ed16-472c-9bb7-fb4b40f7fa9d/
%15   Charadriiformes http://phylopic.org/image/6b77616d-7e99-488a-97e0-c6e919c62ec9/
%16       Galliformes http://phylopic.org/image/2de1c95c-7e1f-429b-9c08-17f0a27d176f/
%17           Diptera http://phylopic.org/image/fdd9c1c9-7391-4e2b-af4e-a0d289fa9466/
%18         Hemiptera http://phylopic.org/image/3010e744-6b06-4022-9ce6-354db93e99fe/

% Accipiter Brisson 1760  Accipitridae  Accipitriformes Savigny 1809  Accipitrimorphae Accipitrinae  Afroaves  Telluraves
% http://phylopic.org/image/f52d11cf-1d3c-41ae-b37e-e46c822a71b3/
% http://creativecommons.org/licenses/by/3.0/
% Shyamal

% http://phylopic.org/image/f68e0e0e-ed16-472c-9bb7-fb4b40f7fa9d/
% http://creativecommons.org/publicdomain/zero/1.0/
% Ferran Sayol